---
title: "COMPSCIX 415.2 Homework 7"
author: "Sumit Pattanayak"
date: "March 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
My Github repository for my assignments can be found at this URL: https://github.com/sumitpattanayak007/compscix-415-2-assignments.git


```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr) 
library(tidyverse)
library(broom)
```

#Exercise 1:
##Load the train.csv dataset into R. How many observations and columns are there?

```{r}
file_path <- "C:/Users/Sumit/Documents/compscix-415-2-assignments/Session7/train.csv"
train_data <- read_csv(file=file_path)
glimpse(train_data)
```

Answer:
Observations: 1,460
Columns: 81

#Exercise 2:
Normally at this point you would spend a few days on EDA, but for this homework we will do some very basic EDA and get right to fitting some linear regression models.

##Q1: Visualize the distribution of SalePrice.
```{r}
train_data %>% 
  ggplot() +
  geom_histogram(aes(SalePrice)) 
```

##Q2: Visualize the covariation between SalePrice and Neighborhood.
```{r}
neighbour_price <- train_data %>% 
  group_by(Neighborhood) %>%
  summarise(SalePriceBar = median(SalePrice, na.rm = TRUE)) %>% 
  arrange(desc(SalePriceBar))
ggplot(neighbour_price) +
  geom_col(aes(Neighborhood, SalePriceBar)) +
  coord_flip()
```

##Q3: Visualize the covariation between SalePrice and OverallQual

```{r}
quality_price <- train_data %>% 
  group_by(OverallQual) %>%
  summarise(SalePriceBar = median(SalePrice, na.rm = TRUE))
ggplot(quality_price) +
  geom_col(aes(OverallQual, SalePriceBar)) +
  coord_flip()
```

#Exercise 3:
 Fit the model and then use the broom package
 
##Q1: take a look at the coefficient
```{r}
(train.lm <- lm(formula = SalePrice ~ OverallQual, data = train_data))
tidy(train.lm)
```
Answer: Coefficient is 45436.

##Q2: compare the coefficient to the average value of SalePrice
```{r}
mean(train_data$SalePrice) 
train.lm
```
Answer: Coeffiicient (45436) is much smaller than Average SalePrice (180921)

##Q3: Take a look at the R-squared

```{r}
glance(train.lm)
```
Answer:  The R-squared is 0.6256519

#Exercise 4:
Now fit a linear regression model using GrLivArea, OverallQual, and Neighborhood as the features.

```{r warning=FALSE}
data(train_data)

ggplot(train_data) +
  geom_point(aes(GrLivArea, SalePrice), alpha = .15) +
  geom_smooth(aes(GrLivArea, SalePrice), method = "lm", se = FALSE)

ggplot(train_data) +
  geom_point(aes(OverallQual, SalePrice), alpha = .15) +
  geom_smooth(aes(OverallQual, SalePrice), method = "lm", se = FALSE)

ggplot(train_data) +
  geom_boxplot(aes(Neighborhood, SalePrice)) +
  coord_flip()
```

##Q1: How would you interpret the coefficients on GrLivArea and OverallQual?

Answer: For every unit of `GrLivArea` increased, the price increases by $55.56. For every unit increase of `OverallQual`, the sale price of the house increases by $20,951.

```{r}
houseprices_lm2 <- lm(formula= SalePrice ~ GrLivArea + OverallQual + Neighborhood, data = train_data)
tidy(houseprices_lm2)
glance(houseprices_lm2)
```

So the feature of OverallQual holds more weight over price increases than GrLivArea.

##Q2: How would you interpret the coefficient on NeighborhoodBrkSide?
Answer: A house in the neighborhood of Brookside, holding square footage and house quality constant, will cost about $13,025 less on average.

##Q3: Are the features significant?
Answer: The features of `OverallQual` and `GrLivArea` are significant values based on the p value. But only a handful of neighborhoods are, favoring the highest median priced neighborhoods. 

##Q4. Are the features practically significant?
Answer: It would seem that looking at `OverallQual` and `GrLivArea` would be conclusive indicators of correlation given the p value but one would have to comb the neighborhood data more diligently as many of the neighborhoods are not good indicators of sales price. 

##Q5. Is the model a good fit?
Answer: Because the adjusted R squared value is close to 1, the model is a good fit. 


#Exercise 6:
One downside of the linear model is that it is sensitive to unusual values because the distance incorporates a squared term. Fit a linear model to the simulated data below (use y as the target and x as the feature), and look at the resulting coefficients and R-squared. Rerun it about 5-6 times to generate different simulated datasets. What do you notice about the model’s coefficient on x and the R-squared values?

```{r}
sim1a <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
)
sim1a
(sim.lm <- lm(formula = y ~ x, data = sim1a))
glance(sim.lm)
ggplot(sim1a) +
  geom_jitter(aes(x,y)) +
  geom_smooth(aes(x, y), method = "lm", se = FALSE)
```

Answer: By looking at the coefficient on x, it is mostly stable, staying within the 1-2 range. The R squared value however would sometimes get smaller. Looking at a plot, one could see that when there is an outlying value, that will distort the R squared value significantly, making it seem like the model is not a good fit. 

