---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Sumit Pattanayak"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
My Github repository for my assignments can be found at this URL: https://github.com/sumitpattanayak007/compscix-415-2-assignments.git


##The tidyverse packages
##---------------------------------------------------------------------------------
##Q1.Can you name which package is associated with each task below? 
Plotting 
Data munging/wrangling 
Reshaping (speading and gathering) data 
Importing/exporting data 

##Answer:
Plotting - ggplot
Data munging/wrangling - tidyr
Reshaping (speading and gathering) data - tidyr
Importing/exporting data - readr in tidyverse

##Q2.Now can you name two functions that you’ve used from each package that you listed above for these tasks?

##Answer:
Plotting - ggplot - geom_boxplot & geom_point
Data munging/wrangling - tidyr - count & mutate
Reshaping (speading and gathering) data - tidyr - gather & spread
Importing/exporting data - readr in tidyverse - read_xxx & write_xxx

##R Basics
##---------------------------------------------------------------------------------
##Q1. Fix this code with the fewest number of changes possible so it works:
My_data.name___is.too00ooLong! <- c( 1 , 2 , 3 )

##Answer:
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
My_data.name___is.too00ooLong
```

##Q2. Fix this code so it works:
my_string <- C('has', 'an', 'error', 'in', 'it)

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

##Q3.  Look at the code below and comment on what happened to the values in the vector. 
my_vector <- c(1, 2, '3', '4', 5) my_vector
 [1] "1" "2" "3" "4" "5"
 
 Answer: All the values are converted to Character by c().
 
##Data import/export (3 points)
##---------------------------------------------------------------------------------
##Q1: Download the rail_trail.txt ﬁle from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.
```{r}
library(tidyverse)
file_path <- "C:/Users/Sumit/Documents/compscix-415-2-assignments/COMPSCIX 415.2 Homework 5/Midterm/rail_trail.txt"
rail_trail <- read_delim(file = file_path, delim = "|")
```
```{r}
glimpse(rail_trail)
```

##Q2. Export the ﬁle into a comma-separated ﬁle and name it “rail_trail.csv”. Make sure you deﬁne the path correctly so that you know where it gets saved. Then reload the ﬁle. Include your export and import code and take another glimpse
```{r}
write.csv(rail_trail, file = "rail_trail.csv")

file_path <- "C:/Users/Sumit/Documents/compscix-415-2-assignments/COMPSCIX 415.2 Homework 5/Midterm/rail_trail.csv"
csv_data <- read_csv(file = file_path)

glimpse(csv_data)
```

##Visualization:
##---------------------------------------------------------------------------------
##Q1. Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.

##Answer:
Dataset for each category is incomplete.
Comparative graph would have been better.
Categories with gender would be more prominent.

##Q2. Reproduce this graphic using the diamonds data set.
```{r}
library(tidyverse)
ggplot(data = diamonds, mapping = aes(factor(cut), carat, fill =  color)) + 
  geom_boxplot(position = "identity") +
  xlab ("CARAT OF DIAMOND") +
  ylab ("CUT OF DIAMOND") +
  coord_flip()
```


##Q3:The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again

```{r}
library(tidyverse)
ggplot(data = diamonds, mapping = aes(factor(cut), carat, fill =  color)) + 
  geom_boxplot() +
  xlab ("CARAT OF DIAMOND") +
  ylab ("CUT OF DIAMOND") +
  coord_flip()
```

##Data munging and wrangling
##---------------------------------------------------------------------------------
##Q1.1. Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy.
```{r}
table2
```

##Answer: The data is not Tidy. Here is my try to make it tidy.
```{r}
table2 %>%
    spread(key = type, value = count)
```
##Q2. Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.

##Answer:
```{r}
transmute (diamonds,
  price_per_carat = price/carat)
```

##Q3. For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.

```{r}
diamonds %>%
   filter(carat < 1.5 & price > 10000) %>%
   count(cut) %>%
   mutate(prop = n / sum(n)) %>%
   print(n = 30)
```

• Do the results make sense? Why? 
The reult does make sense as all the number in each catogory are increasing as per the cut.

• Do we need to be wary of any of these numbers? Why?
Categories of Diamond are not consistant. The price of Very Good and Premium are pretty close. 

## EDA
##---------------------------------------------------------------------------------
Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:
```{r}
glimpse(txhousing)
```

#Q1: During what time period is this data from? 
Answer: The data is from 2000 to 2015.

#Q2: How many cities are represented?
```{r}
txhousing %>%
  count(city)
```
Answer: There are 46 cities.

#Q3. Which city, month and year had the highest number of sales?
```{r}
txhousing %>%
  select(city,month,year,sales) %>%
  group_by(city) %>%
  filter(sales == max(sales))
```
##Answer:
Houston has the highest sale in the month of July of 2015.

#Q4. What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work

#Q5. What proportion of sales is missing for each city?
```{r}
txhousing %>%
   filter(is.na(sales)) %>%
   count(city) %>%
   mutate(prop = n / sum(n)) %>%
   print(n = 30)
```

#Q6. Looking at only the cities and months with greater than 500 sales:
```{r}
txhousing %>%
   filter(sales > 500) %>%
   select(city, month, sales)
```

• Are the distributions of the median sales price (column name median), when grouped by city, diﬀerent? The same? Show your work. 

```{r}
txhousing %>%
  filter(sales > 500) %>%
  select(city, month, sales) %>%
  group_by(city) %>%
  summarise(
    median = median(sales))
```
Answer: They are different.

• Any cities that stand out that you’d want to investigate further? 

Answer: Not in general.

• Why might we want to ﬁlter out all cities and months with sales less than 500?
Answer: Not sure. 