---
title: "COMPSCIX 415.2 Homework 1"
author: "Sumit Pattanayak"
date: "February 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
My Github repository for my assignments can be found at this URL: https://github.com/sumitpattanayak007/compscix-415-2-assignments.git

```{r}
library(mdsr)
library(tidyverse)
```

```{r}
data("WorldCities")
###### Here is a glimpse of dataset:
glimpse(WorldCities)
```

####Extract parts of the data

```{r}
WorldCities <- head(WorldCities, 200)
country_col <- WorldCities$country
unique(country_col)

```

###QUESTION 2: There is a column called region in the data set. Can you extract this and show only the unique values?

```{r}
reg <- WorldCities$region
unique(reg)

```

##QUESTION 3: Can you extract and show the unique entries from the country column in WorldCities using one line of code and two %>% operators? The output will look like this:

```{r}
WorldCities %>% select(country) %>% unique()

```

```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
```

###QUESTION 4: Make one more histogram of the population and add a new x-label, a new ylabel, a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().

```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + xlab('Population') + ylab('Freqency') + ggtitle('Histogram of Population of World Cities') + theme_bw()
```

