---
title: "COMPSCIX 415.2 Homework 2"
author: "Sumit Pattanayak"
date: "February 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 3.2.4

```{r}
library(ggplot2)
1
ggplot(data = mpg)
```
Output : Nothing is showing here, blank.


2

```{r}
nrow(mpg)
ncol(mpg)
```

Output : No of Rows - 234
Output : No of Colms - 11

3
f = front-wheel drive, r = rear wheel drive, 4 = 4wd
f indicates whether the vehicle is front-wheel drive or rea wheel drive or 4 wheel drive


4
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```


5
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

Points on the plot are overlapping with each other

## Exercise 3.3.1
## 1

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```
## Blue dots are showing Red as color argument was set within aes(), not geom_point()
## Correct code would be like below:

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

## 2
mpg
## Categorical - manufacturer, model, trans, drv, fl, class
## Continuous - displ, cyl, cty, hwy
## Categorial - <chr>
## Continuous - <dbl> and <int>


## 3 - Map a continuous variable to color, size, and shape

## using color
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))

## using size
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))

## using shape
## ggplot(data = mpg) +
##  geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
## Error: A continuous variable can not be mapped to shape

##continuous variables are visualized on a spectrum
##categorical variables are binned into discrete categories

## 4 - What happens if you map the same variable to multiple aesthetics?
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))

## Both aesthetics are implemented, but multiple legends are generated.

## 5 - What does the stroke aesthetic do? What shapes does it work with?  
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), stroke = 3, shape = 21)
## stroke adjusts the thickness of the border for shapes that can take on different colors both inside and outside.

##6 - What happens if you map an aesthetic to something other than a variable name, like  aes(colour = displ < 5)?

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))

## displ is temporary variable to hold the result whether True or False
```

## Exercise 3.5.1
```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

## 4
## Advantage - Faceting splits the data into separate grids and better visualizes trends within each individual facet.
## Disadvantage - harder to visualize the overall relationship across facets
## The color aesthetic is fine when your dataset is small, but with larger datasets points may begin to overlap with one another. In this situation with a colored plot, jittering may not be sufficient because of the additional color aesthetic.

##5
?facet_wrap
## nrow, ncol	- Number of rows and columns.
## as.table	- If TRUE, the default, the facets are laid out like a table with highest values at the bottom-right. If FALSE, the facets are laid out like a plot with the highest value at the top-right.
## in `facet_grid` nrow and ncol are implied by the faceting variables. `dir` also controls the placement of the individual panels, and so isn't an argument of `facet_grid`.

```

## Exercise 3.6.1
```{r}
library(ggplot2)
## 1 - What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
## Line chart - geom_line()
## Boxplot - geom_boxplot()
## Histogram - geom_histogram()
## Area chart - geom_area()

## 2
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

## 3 - What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
## It removes the legend. 
## The aesthetics are still mapped and plotted, but the key is removed from the graph. 
## I don’t know why used it earlier because he actually did not.

## 4 - What does the se argument to geom_smooth() do?
## It determines whether or not to draw a confidence interval around the smoothing line.

## 5 - Will these two graphs look different? Why/why not?
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

## Answer -No because they use the same data and mapping settings. The only difference is that by storing it in the  ggplot() function, it is automatically reused for each layer.

## 6 - Recreate the R code necessary to generate the following graphs.
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(aes(group = drv), se = FALSE) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4, colour = "white") + 
  geom_point(aes(colour = drv))
```

## 3.7.1

## 2 - What does geom_col() do? How is it different to geom_bar()?
## `geom_col` leaves the data as it is. `geom_bar()` creates two variables (count and prop) and then graphs the count data on the y axis. With `geom_col` you can plot the values of any x variable against any y variable.
