---
title: "COMPSCIX 415.2 Homework 9/Final"
author: "Sumit Pattanayak"
date: "April 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(broom)
```

#Exercise 1:
## Sampling Distributions, Functions and For Loops

###You can sample from this exponential distribution in R using this code:
```{r}
# sample size
samp_size <- 100
# set the rate parameter
samp_rate <- 1/10000
# take sample
rexp(n = samp_size, rate = samp_rate)
```


##STEP 1
*Write an R function that does the following:*

*Takes a sample of size samp_size from this exponential distribution (samp_size is an input parameter for the function)*
*Calculates the mean of that sample*
*Calculates the standard deviation of that sample*
*Returns the calculated mean and standard deviation as a list*

```{r}
samp_fun <- function(samp_size, samp_rate) {
  exp_samp <- rexp(n = samp_size, rate = samp_rate)
  samp_avg = mean(exp_samp)
  samp_std_dev = sd(exp_samp)
  stats <- list(samp_avg = samp_avg, samp_std_dev = samp_std_dev)
return(stats)
}
```

##STEP 2
*Then write a loop that does this:*
*• Runs the above function 1000 times, with samp_size = 50 and samp_rate = 1/10000*
*• Saves all of the sample means in a vector called sample_means, and all of the sample standard deviations in a vector called sample_sds*
```{r}
n <- 1000
samp_size = 50

main_function <- function(n,samp_size){
#empty vector for saving sample means and SDs
  sample_means <- rep(NA, n)
  sample_sds <- rep(NA, n)

  for(i in 1:n) {
    x <- samp_fun(samp_size, 1/10000) # save sample mean and sd
    sample_means[i] <- x$samp_avg # add sample mean to vector
    sample_sds[i] <- x$samp_std_dev # add sample sd to vector
  }

# save output in a tibble
  exp_samp_df <- tibble(
    sample_means, 
    sample_sds
  )
return(exp_samp_df)
}
main_function(n,samp_size)
```

##STEP 3
*Then*
*• plot your sample means as a histogram*
*• output the standard deviation of your sample means*
*• calculate the theoretical standard error (σ = 10000, n = sample size)*
*• calculate the mean of the sample standard deviations and use this to calculate the empirical standard error*
*Plot the sample means*
```{r}
  exp_samp_df <- main_function(n,samp_size)
  plot <- exp_samp_df %>% ggplot(aes(x = sample_means)) +
  geom_histogram()
  plot
# Standard deviation of the sample means
  sd_sample_means <- sd(exp_samp_df$sample_means)
  sd_sample_means
# Calculate the theoretical standard error
  theo_sd_error <- 10000/sqrt(samp_size)
  theo_sd_error
# Calculate the mean of the sample standard deviations and use this to calculate the   empirical standard error
  mean_sample_sd <- mean(exp_samp_df$sample_sds)/sqrt(samp_size)
  mean_sample_sd
```

##STEP 4
*Repeat STEP 2 and STEP 3 using a sample size of 5000*


### for Sample size 5000
```{r}
#Step 2
samp_size = 5000
main_function(n,samp_size)
#Step 3
  exp_samp_df <- main_function(n,samp_size)
  plot <- exp_samp_df %>% ggplot(aes(x = sample_means)) +
  geom_histogram()
  plot
# Standard deviation of the sample means
  sd_sample_means <- sd(exp_samp_df$sample_means)
  sd_sample_means
# Calculate the theoretical standard error
  theo_sd_error <- 10000/sqrt(samp_size)
  theo_sd_error
# Calculate the mean of the sample standard deviations and use this to calculate the   empirical standard error
  mean_sample_sd <- mean(exp_samp_df$sample_sds)/sqrt(samp_size)
  mean_sample_sd
```

# Exercise 2:
##Linear Regression

```{r}
# Load the train.csv dataset
train_df_path <- 'C:/Users/Sumit/Documents/compscix-415-2-assignments/Final/train.csv'
train_df <- read_csv(file = train_df_path)

# Fit a regression model with y = SalePrice and use LotArea, OverallQual, and ExterQual as features
sales_mult_lm <- lm(formula = SalePrice ~ LotArea + OverallQual + ExterQual, data = train_df)
```

*Answer these questions:*

1. Use the broom package to output the coefficients and the R-squared
```{r}
library(broom)
tidy(sales_mult_lm)
glance(sales_mult_lm)
```

2. Interpret the coefficient on LotArea

Answer: For every one unit increase in LotArea, the sale price increases, on average, by $1.45.

3. Interpret the coefficient on ExterQualGd

Answer: ExterQual is an unordered categorical variable, so the coefficient can be interpreted as a comparison to the first value in ExterQual. Therefore, the mean price difference between ExterQualGd and the first value (ExterQualEx) is -$71,529.

4. Compare this model to the model we fit in HW 7 with GrLivArea, OverallQual, Neighborhood. Which is the better fitting model?

Answer: The model we fit in HW 7 with GrLivArea, OverallQual, Neighborhood had an adjusted R-squared value of 0.78. Our model here does not fit as well, shown by the lower adjusted R-squared value of 0.69. Therefore the model fit for HW 7 is better.

#Exercise 3:
##AB Testing


*Answer these questions:*

1. What proportion of visitors converted for each version of the webpage?
```{r}
# Load the ab_test_data.csv dataset
ab_test_df_path <- 'C:/Users/Sumit/Documents/compscix-415-2-assignments/Final/ab_test_data.csv'
ab_test_df <- read_csv(file = ab_test_df_path)

ab_test_df$conversion <- as.character(ab_test_df$conversion)

ab_freq_df <- ab_test_df %>% count(version, conversion)

ab_freq_df <- ab_freq_df %>% 
  spread(key = conversion, value = n) %>%
  rename(not_converted = '0', converted = '1') %>%
  mutate(total = not_converted + converted,
         prop_converted = converted/total,
         perc_converted = paste0(round(prop_converted*100, 1), "%"))

ab_freq_df
```
Answer: 4.2 percent of visitors converted for version A and 10 percent of visitors converted for version B.

2. Perform the AB test in R. What is the p-value for the AB test (hypothesis test of proportions)?

```{r}
samp_a <- ab_freq_df %>% filter(version == "A") %>% select(converted)
samp_a <- samp_a[[1]]
samp_b <- ab_freq_df %>% filter(version == "B") %>% select(converted)
samp_b <- samp_b[[1]]
n_a = n_b = 2000

two_prop_test <- prop.test(c(samp_a, samp_b), c(n_a, n_b))
two_prop_test$p.value
```

Answer: The p-value is very small (less than 0.000), which tells us that the conversion rates for Version A and B are significantly different than each other.